### 2021-04-23 Modules/abstractEnglish_LOCAL.tex
Before final comments

% State-of-the-art array processing methods, ranging from high-order statistics to adaptive configuration, require costly computing efforts in pursuit for spatial performance improvement.
The general field of array processing has been thoroughly studied throughout several decades.
The array sensors' spatial diversity enables the extraction of spatial information about impinging signals, thus laying the ground for wide range of applications.
The array size and the number of its elements ($N$) has significant influence on the obtained array performance, such as SNR improvement, spatial separation capabilities and its spatial response's degrees of freedom (DOF).
\par 
Considering localization related applications, several processing approaches are to be found. The most basic approach is \emph{beamforming}, i.e steering the array to different parts of the arena in search of objects.
Another approach, which is probably the most commonly used, is the subspace-based processing which tries to identify input signals that fit into the array's manifold using the orthogonal projection concept.
Statistical processing, although costly in terms of computation effort, is also in use, ranging from maximum-likelihood (ML) related schemes to higher-order-statistics (HOS) using cumulants.
Newer approaches which are still in research stage use neural networking and sparsity-based algorithms.
\par
Although much researched, in this work we revisit the \emph{beamforming} approach.
ULA based beamforming and temporal finite impulse response (FIR) filtering are mathematically analogous, where the DOA acts as the spatial version of temporal frequency.
Inspired by this analogy, we search for the spatial counterpart of the temporal infinite impulse response (IIR) filter.
To this end, we arbitrarily choose to use localization related problem formulation and suggest a feedback based approach, featuring low complexity and high spatial performance in the mere excess of integrating a transmitter to the array.
Considering RADAR-like arrays, the transmitter is already in place, hence a mere processing modification suffices.
Assuming the target of interest has a mirror-like behaviour (i.e., reflects its impinging signals), the spatial feedback between the array and the target is created by continuously re-transmitting a synthesized version of the impinging signal (and its reflections) to the target.
Hence a spatial loop is created, which is shown to be equivalent of enlarging the array's aperture in terms of spatial resolution.
Also notable is that the achievement of the exclusive IIR-like beampattern is purely done in the spatial domain while avoiding any temporal processing to the signal.
\par 
Using a traditional beamforming performance analysis, the beamwidth, peak to side-lobe ratio, array directivity and white noise sensitivity are evaluated for the feedback based array.
A significant improvement in all aspects is shown, while thoroughly discussing the conditions for enhanced performance.
Considering ideal scenarios, the feedback beamformer virtually achieves an infinite aperture, increasing the available spatial information about the target and significantly improves the array's spatial performance.
\par
Taking into account the unavoidable estimation errors and uncertainties, we find that the basic feedback integrated beamformer is very sensitive to even mild range errors.
% Also, the system is more sensitive as the carrier frequency increments.
% Considering practical carrier frequencies, the range sensitivity is of wavelength order. 
% Seeking for a solution, bearing in mind that summing multiple harmonics produce also other lower / higher frequencies, we propose a practical and robust implementation of the feedback-based localization concept.
As a solution, we propose a more complex architecture, using two harmonics, which we call \emph{dual-frequency feedback beamformer}.
We also thoroughly analyse the proposed solution and find that it features low and controllable estimation errors sensitivity in the mere expense of doubling the computation effort.

>>>>

The general field of array processing has been thoroughly studied throughout several decades.
The array sensors' spatial diversity enables the extraction of spatial information about impinging signals, thus laying the ground for wide range of applications.
The array size and the number of its elements have significant influence on the achievable array performance, such as signal to noise ratio (SNR) improvement, spatial separation capabilities, directivity, array gain, source localization, etc.
\par 
Localization applications are divided to several groups, each featuring a different approach for estimating the spatial parameters of objects in the arena.
The most basic approach is \emph{beamforming}, where the arena (or a part of it) is continuously scanned by steering (mechanically or electronically) the array to all possible directions in search for objects.
Another approach, more commonly used, is the subspace-based processing which exploits the orthogonality between the array manifold (i.e. the array's excitations set of impinging planewaves from all directions) and the noise subspace.
In this scheme the input signal is projected onto the array manifold - thus mitigating non directional noise interference. 
Statistical processing, although costly in terms of computation effort, is also in use.
One important example is the maximum-likelihood (ML) approach which assumes a known probability density function (PDF) for the input signal or the noise.
\par
In this work we revisit the \emph{beamforming} approach.
Inspired by the analogy between spatial processing with uniform linear array (ULA) and temporal finite impulse response (FIR) filtering, we search for the spatial counterpart of the temporal infinite impulse response (IIR) filter.
To this end, we formulate a localization problem and suggest a feedback based approach, featuring low complexity and high spatial performance in the mere excess of integrating a transmitter to the array.
Considering RADAR-like arrays, the transmitter is already in place, hence a mere processing modification suffices.
\par  
Assuming the target of interest has a mirror-like behaviour (i.e., reflects its impinging signals), the spatial feedback between the array and the target is created by continuously re-transmitting a synthesized version of the impinging signal (and its reflections) to the target.
In this manner, a spatial loop is created, which is shown to be equivalent of enlarging the array's aperture in terms of spatial resolution.
\par 
Using a traditional beamforming performance analysis, the beamwidth, peak to side-lobe ratio, array directivity and white noise sensitivity are evaluated for the feedback based array.
A significant improvement in all aspects is shown, while thoroughly discussing the conditions for enhanced performance.
Considering ideal scenarios, the feedback beamformer virtually achieves an infinite aperture, increasing the available spatial information about the target and significantly improves the array's spatial performance.
\par
Taking into account the unavoidable estimation errors and uncertainties, we find that the basic feedback integrated beamformer is very sensitive to even mild range errors.
% Also, the system is more sensitive as the carrier frequency increments.
% Considering practical carrier frequencies, the range sensitivity is of wavelength order. 
% Seeking for a solution, bearing in mind that summing multiple harmonics produce also other lower / higher frequencies, we propose a practical and robust implementation of the feedback-based localization concept.
As a solution, we propose a more complex architecture, using two harmonics, which we call \emph{dual-frequency feedback beamformer}.
We also thoroughly analyse the proposed solution and find that it features low and controllable estimation errors sensitivity in the mere expense of doubling the computation effort.

### 2021-04-23 main/intro.tex
Before final comments

@@

MUSIC (Multiple SIgnal Classification) algorithm \cite{schmidt1986multiple} which decomposes the estimated covariance matrix to its signal and noise eigenvectors, using orthogonal projection exploiting the fact that the noise space is orthogonal to the signal subspace.

@@

Though this approach served as a basis for extensive research, leading to more advanced concepts such as the ICA \cite{hyvarinen1999survey}, in the context of practical DOA estimation it's computational effort is high.

@@

% The main advantage of the first (i.e. virtual array) over the latter (i.e. minimum redundancy) is the obtainable analytical expression for the improvement in spatial selectivity when using virtual arrays.
% In the minimum redundancy approach no such expression is available.
% On the other side, the computation effort is considerably higher when the 2-nd order statistics are computed.
Although the $q$-th order statistics based approach promises substantial improvement, both noise sensitivity and impractical computation amount are dominant drawbacks, limiting the usage of such methods in practical applications.   

@@

As in many adaptive processing applications, the two methods rely on steepest descend optimization, which exposes, the adaptive processing approach to errors related to optimization parameters choices such step size etc.

@@

@@

To this end, we arbitrarily choose to formulate the problem in the context of a localization problem, hence our goal is to estimate the direction and the range of some target of interest. 

@@

\par In the followings, we elaborate in Chapter~\ref{chap:prelims} on some relevant theoretic concepts which serve as the foundation of this work.
Then, our novel architecture is presented and analysed in Chapter.~\ref{chap:firstchap} followed by a thorough concluding discussion of the findings and future research directions in Chapter.~\ref{chap:conclusion}.

>>>>

\section{Background and Motivation}

@@

MUSIC (Multiple SIgnal Classification) algorithm \cite{schmidt1986multiple} which decomposes the estimated covariance matrix to its signal and noise eigenvectors, using orthogonal projection exploiting the fact that the noise sub space is orthogonal to the signal subspace.

@@

Though this approach served as a basis for extensive research, leading to more advanced concepts such as the independent component analysis (ICA) \cite{hyvarinen1999survey}, in the context of practical DOA estimation it's computational effort is high.

@@

Although the $q$-th order statistics based approach promises substantial improvement, both noise sensitivity and impractical computation costs are dominant drawbacks, limiting the usage of such methods in practical applications.   

@@

As in many adaptive processing applications, the two methods rely on steepest descend optimization, which exposes the adaptive processing approach to errors related to optimization parameters choices such step size etc.

@@

\section{Overview of the Thesis}
In this thesis, we propose a new feedback based beamformer architecture which acts as an actual spatial IIR filter while avoiding any temporal processing to the signal.
We aim to guide the reader thorough the key stages of the research where in each stage, a sufficient background is provided followed by the solution composition.
\par As this research does not closely follows any former work, some of analysis must be rigorously justified, hence we turn to classic literature as guidance.
As an example, for the configuration of the array, we use the Fisher Information Matrix (FIM) aiming to maximize the residing spatial information in the system.
\par After presenting the basic approach with it's complementary adjusted classic spatial analysis, we find that it is inherently sensitive to range estimation errors.
We then suggest a more robust solution which is based on our conclusions from previous analysis.
With the robust solution in hand, we present simulations and compare it to the basic conventional beamformer and the initial more sensitive approach.
As expected, we find that it outperforms them, proving to perform well also in low signal to noise ratios (SNR).
\section{Main Contributions}

@@

To this end, we arbitrarily choose to formulate the problem in the context of localization, hence our goal is to estimate the direction and the range of some target of interest. 

@@

\section{Thesis Organization}
\par This thesis is organized as follows.
In Chapter~\ref{chap:prelims}, we cover basic concepts of array processing, followed by some emphasize on the localization field of research and finalize with the basic concepts which drive the motivation for this research.
\par In Chapter.~\ref{chap:firstchap} the basic feedback beamformer is presented and thoroughly analysed, followed by the presentation of the more robust architecture, which uses the former as a building block.
Simulations are also provided, supporting the analytical analysis.
\par In Chapter.~\ref{chap:conclusion} we discuss the results and conclude our findings. We finalize by also proposing some leads to future research.
